Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-11-28 src/tex/ms.tex
Update ms.tex

\section{Introduction}
\label{sec:Introduction}
% Motivation
Since Einstein proposed his theory of general relativity (GR), it was tested in
a wide range of length scales. After a century, we know that GR does not agree
with quantum theories at some length scale. To unify both theories, we need to
study the possibility of different beyond-GR theories. \kw{I think now it kind
of goes into birefringence too fast. Maybe talk about what effect one may expect
to see in GW in general a bit? Say there is polarization modification, phasing
modification, etc.} Some beyond-GR theories such as Chern-Simons gravity
suggests that there is gravitational wave (GW) amplitude birefringence, while GR
predicts there is no birefringence. GW amplitude birefringence is a property of
space-time which consists of the enhancement of one GW polarization over the
other during the propagation of the wave. the local modification of the
amplitude is small, but it grows with the number of cycles the GW is propagated.

% Previous studies
Previous studies have constrained amplitude birefringence by performing different statistical analysis.
In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2,
the second GW transient catalog, and use the distribution to constrain the
amplitude birefringence effect with the assumption of frequency independence in
GW amplitude birefringence. In \citet{Yamada_2020} and \citet{Wang_2021}, they
both modeled the waveform under birefringence effect with a simplify assumption
\kw{What is the simplify assumption} and performed parameter estimation (PE) on the
events in GWTC-1, the first GW transient catalog.

% What's new?
In this paper, we used a birefringence model with higher order terms which
include the frequency dependence of amplitude birefringence to perform a more
accurate PE \kw{I think there is qualitative difference
between this work and former works, so it is not just more accurate}. We also
performed Bayesian inference on events in GWTC-3, the third GW transient
catalog. We included 69 binary black hole merger events with smallest false
alarm rate $\leq1\mathrm{yr^{-1}}$. These events are also listed in TABLE I in
\citet{GWTC_3_population}.\footnote{GW190720 and GW200129 are included in the
list in \citet{GWTC_3_population}, but we did not include them in this study.
Details in Discussion.}

@@

% Parameter estimation
The major objective of this study is to constrain GW amplitude birefringence,
which is represented by the opacity parameter $\kappa$. To constrain $\kappa$,
we perform PE with data from the third GW transient catalog \citep{GWTC-2.1,
GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which is
able to calculate posteriors of GW parameters based on interferometer data and
priors for the parameters in question \citep{Bilby}. We set the prior on
$\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using
the modified waveforms. In both case, we use the same PE configuration as in
\citet{GWTC-2.1, GWTC-3} for each individual events. We also perform test to
check if the PE could recover the PE done by LIGO-Virgo-KAGRA Collaboration
\kw{State that this is a sanity check} (LVK) in \citet{GWTC-2.1, GWTC-3}. We set
the prior on $\kappa$ to be a $\delta$ function at $0$, which is the same as
using GR \kw{This sentence kind of read odd. Add conjunction to link it to the
GR test.}. These control tests result can be found in the \href{https://zenodo.org/record/7338924}{data release}.

@@

% How to calculate the combined posterior
In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$
of each events obtained from PE. \kw{Rewrite the following paragraph in the
following way: 1. There are a set of measured $\kappa$ 2. State that there can
either be a single true $\kappa$ or a distribution of $\kappa$ 3. Describe
hierarchical Bayesian, which is essentially stacking the individual results
together. 4. Your choice of model and reason behind such choice.} We assume
$\kappa$ is different for each event, and the population distribution of
$\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation
$\sigma$. We then performed hierarchical modeling \kw{Need to describe what
hierarchical modelling is} to construct the likelihood function of $\mu$ and
$\sigma$ given interferometer data. To sample the posterior distribution of
$\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC},
with prior of $\mu$ and $\sigma$ being uniform distributions. The posterior of
$\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}. The mean of
$\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.

@@

% Case: GW150914 (example of broken degeneracy)
\subsection{Result on GW150914}
Some events have bimodal $\kappa$ distribution, but before we show a bimodal
event, we should understand the mechanism of a unimodal $\kappa$ distribution.
\kw{Just say what you are trying to say, instead of leaving a cliff hanger}
Consider GW150914, the first GW detected by LIGO. In figure
\ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
With the frequency independent birefringence model, the posteriors for
$\cos\iota$ look different from the posteriors assuming GR. This is because, for
a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if
the frequency dependence is not included. To reconstruct the amplitude ratio
from the interferometer data, an $\iota$ representing a more face-off inspiral
can pair with a positive $\kappa$, or an $\iota$ representing a more face-on
inspiral with a negative $\kappa$. As a result, many pairs of $\iota$ and
$\kappa$ can be plausible with the frequency independence. \kw{Many pairs
doesn't mean bimodelity, you should say why the middle part is constrained to be small as well.}

>>>>

\section{Introduction}
\label{sec:Introduction}
% Motivation
Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR does not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
\kw{I think now it kind of goes into birefringence too fast.
Maybe talk about what effect one may expect to see in GW in general a bit?
Say there is polarization modification, phasing modification, etc.}
Some beyond-GR theories such as Chern-Simons gravity suggests that there is gravitational wave (GW) amplitude birefringence, while GR predicts there is no birefringence.
GW amplitude birefringence is a property of space-time which consists of the enhancement of one GW polarization over the other during the propagation of the wave.
The local modification of the amplitude is small, but it grows with the number of cycles the GW is propagated.

% Previous studies
Previous studies have constrained amplitude birefringence by performing different statistical analysis.
In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption.
They assumed the modification of amplitude birefringence is small and performed parameter estimation (PE) on the events in GWTC-1, the first GW transient catalog.

% What's new?
In this paper, we used a birefringence model with higher order terms which include the frequency dependence of amplitude birefringence to perform a more accurate PE.
\kw{I think there is qualitative difference between this work and former works, so it is not just more accurate}
We also performed Bayesian inference on events in GWTC-3, the third GW transient catalog.
We included 69 binary black hole merger events with smallest false alarm rate $\leq1\mathrm{yr^{-1}}$.
These events are also listed in TABLE I in \citet{GWTC_3_population}.
~\footnote{GW190720 and GW200129 are included in the list in \citet{GWTC_3_population}, but we did not include them in this study.
Details in Discussion.}

@@

% Parameter estimation
The major objective of this study is to constrain GW amplitude birefringence, which is represented by the opacity parameter $\kappa$.
To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby, a Bayesian toolkit for GW data analysis which is able to calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question \citep{Bilby}. \
We set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$, and perform PE using the modified waveforms.
We use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each individual events.
We also perform sanity test to check if our PE could recover the PE done by LIGO-Virgo-KAGRA Collaboration (LVK) in \citet{GWTC-2.1, GWTC-3}.
To implement the sanity check, we need to assume there is no birefringence by setting the prior on $\kappa$ to be a $\delta$ function at $0$, which means GR is recovered.
These sanity tests result can be found in the \href{https://zenodo.org/record/7338924}{data release}.

@@

% How to calculate the combined posterior
In figure \ref{fig:kappa_stacked}, we show the individual posterior of $\kappa$
of each events obtained from PE.

\kw{Rewrite the following paragraph in the following way:
1. There are a set of measured $\kappa$
2. State that there can either be a single true $\kappa$ or a distribution of $\kappa$
3. Describe hierarchical Bayesian, which is essentially stacking the individual results together.
4. Your choice of model and reason behind such choice.}

We assume $\kappa$ is different for each event, and the population distribution of $\kappa$ is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
We then performed hierarchical modeling
\kw{Need to describe what hierarchical modelling is}
to construct the likelihood function of $\mu$ and $\sigma$ given interferometer data.
To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}, with prior of $\mu$ and $\sigma$ being uniform distributions.
The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
The mean of $\mu$ and $\sigma$ are \variable{output/Gaussian_mean.txt}, respectively.

@@

% Case: GW150914 (example of broken degeneracy)
\subsection{Result on GW150914}
Some events have bimodal $\kappa$ distribution, but before we show a bimodalcevent, we should understand the mechanism of a unimodal $\kappa$ distribution.
\kw{Just say what you are trying to say, instead of leaving a cliff hanger}
Consider GW150914, the first GW detected by LIGO.
In figure \ref{fig:corner_GW150914}, we show the posteriors of the parameters of GW150914.
With the frequency independent birefringence model, the posteriors for $\cos\iota$ look different from the posteriors assuming GR.
This is because, for a nonprecessing system, there is a degeneracy between $\kappa$ and $\iota$ if the frequency dependence is not included.
To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
\kw{Many pairs doesn't mean bimodelity, you should say why the middle part is constrained to be small as well.}
